Search for Spatially Extended Fermi-LAT Sources Using Two Years of Flight Data

### 2011-08-07 extended_catalog/paper/extended_catalog.tex
many improvements

\usepackage[textwidth=0.8in,colorinlistoftodos]{todonotes}

@@

% some units
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\renewcommand{\sec}{\text{s}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_\text{ext}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_\text{inc}}\xspace}
\newcommand{\loglikelihood}{\ensuremath{LL}\xspace}

@@

\newcommand{\tsextpointlike}{\ensuremath{\tsext_{,\pointlike}}\xspace}
\newcommand{\tsextgtlike}{\ensuremath{\tsext_{,\gtlike}}\xspace}
\newcommand{\tsextalt}{\ensuremath{\tsext_{,\alt}}\xspace}

\newcommand{\ts}{\text{TS}\xspace}
\newcommand{\glon}{\text{GLON}\xspace}
\newcommand{\glat}{\text{GLAT}\xspace}
\newcommand{\alt}{\text{alt}\xspace}

@@

% the program names
\newcommand{\pointlike}{\text{\em pointlike}\xspace}
\newcommand{\python}{\text{\em python}\xspace}
\newcommand{\gtlike}{\text{\em gtlike}\xspace}
\newcommand{\gtobssim}{\text{\em gtobssim}\xspace}
\newcommand{\minuit}{\text{\em Minuit}\xspace}

@@

The Large Area Telescope (LAT) is a pair conversion telescope on the
Fermi Gamma-Ray Space Telescope (Fermi). It has been surveying the
$\gamma$-ray sky since June 2008.  Fermi is well suited as a survey
instrument because of its broad energy coverage (20 \mev to $>300$
\gev), wide field of view ($\sim 2.4 \sr$), and large effective area
($\sim 8000 \cm^2$ at $>1 \gev$).

Using two years of all sky surveying data, the LAT Collaboration
published a catalog of 1873 \gev sources\cite{second_cat}. Many of these can
be positionally associated with a variety of source classes including
Active Galactic Nuclei (AGN), pulsars, galaxies, and supernova
remnants (SNRs).  While many of these classes can be spatially
resolved when observed at other frequencies, detection an extension at
GeV energies by the LAT is complicated by the rather large
point-spread function (PSF) of the instrument (the angular resolution
for events converting in the front part of the detector, defined as
the 68\% containment radius for a point source, is approximately
$3.5^{\circ}$ at 100 MeV, improving to about 0.1$^{\circ}$ at 10 GeV).

@@

In addition to these GeV $\gamma$-ray detections, a variety of
spatially extended sources have been detected in the Galactic plane at
\tev energies using Air Cherenkov detectors. Most prominent was a
survey of the Galactic plane using the High Energy Stereoscopic System
(H.E.S.S) which presented 14 spatially extended sources
(\cite{hess_plane_survey}) sources with extensions varying from
$~0.1\deg$ to $~0.25\deg$. In fact, within our Galaxy only very few
sources (most notably the $\gamma$-ray binaries
LS\,5039 (\cite{HESSLS5039}), LS I+61-303 (\cite{MAGICLSI, VERITASLSI}) and
HESS\,J0632+057 (\cite{HESS0632}) are amongst the few Galactic sources
for which no significant extension was detected at \tev
energies. Given these findings, it can be expected that if the source
population at GeV energies has similar properties, that at least some
of the Fermi-LAT sources in our Galaxy should be spatially extended at
GeV energies.

Being able to spatially resolve the \gev emission from a source is
important for several reasons. Because of the large PSF and the large
candidate source density in our Galaxy, source confusion is a severe
problem for the Fermi-LAT. Finding a matching morphology can be used
to uniquely identify the source.  Because of the strongly varying PSF
of the LAT, the spatial and spectral information about a source do not
simply decouple. Therefore, a biased spatial model will bias the
spectral model of the source and studying the extension of the source
can correct for this.  Furthermore, correctly modeling a source's
extension is important for improving the overall model of the sky and
removing excess residuals.  These residuals manifested themselves in
1FGL as spurious point sources in the region region surrounding the
Large Magellanic Cloud (\cite{first_cat}).  For this reason, we are
motivated to take the 1873 sources that were detected in 2FGL and
apply a search for those sources which are spatially extended.

@@

The convolution of radially symmetric extended sources is optimized by
using a semi-analytic calculation.  The observed photon distribution
can be written as the convolution of the source shape ($I_\text{src}$)
with the psf
\begin{equation}
  \text{PDF}(\vec r) = \int  \text{PSF}(|\vec r - \vec r'|)I_\text{src}(\vec r') d A'.
\end{equation}
For the LAT, the PSF is parameterized by a {\em King function}
\begin{equation}
  \text{PSF}(\vec r) = 
  \frac{1}{2\pi\sigma^2}
  \left(1-\frac{1}{\gamma}\right)
  \left(1+\frac{u}{\gamma}\right)^{-\gamma},
\end{equation}
where $\sigma$ and $\gamma$ are free parameters
(\cite{matthew_kerr_thesis}).  For radially symmetric extended sources,
the angular part of the integral can be done analytically
\begin{equation}
  \text{PDF}(u)= \int_0^\infty dv
  I_\text{src}(v) 
  \left(\frac{\gamma-1}{\gamma}\right)
  \left( \frac{\gamma}{\gamma + u + v}\right)^\gamma 
  \times ~_2F_1 \left(\gamma/2,\frac{1+\gamma}{2},1,\frac{4uv}{(\gamma+u+v)^2}\right).
\end{equation}
This convolution formula is used by \pointlike because it is faster
and more closely reduces to the PSF when the source's size is 
small.

@@

We found that the {\em HESSE} errors on the extension of a source
were sometimes badly estimated.  On the other hand, the {\em Simplex}
algorithm was prohibitively slow and would often fail to convergence. Our
approach to calculating extension errors is to fix the position of the
source and varying just the extension using the {\em Simplex} algorithm.
This approach makes the approximation that the covariance between each
spatial parameter is small but not that the likelihood is Gaussian. This
method is shown schematically in figure~\ref{extension_profile} which
shows the change in \loglikelihood when varying the extension of the
SNR IC443.  The localization error is separately calculated by fixing
the extension and spectrum of the source and fitting a parabola to the
likelihood as a function of position.

@@

The half $\chi^2_1$ distribution was found to hold for the special case
of source finding using EGRET (\cite{mattox_egret}). This distribution
is also plausible when testing point sources for extension. Due to
statistical fluctuations, half of the point sources might look narrower
than the PSF leading to $\tsext=0$. The other half of the time, we would
get a distribution consistent with Wilk's theorem. We will present a
Monte Carlo study that shows an empirical distribution of \tsext in the
null hypothesis and compares it to this distribution.

We simulated point sources of varying spectral models and fit them
them using \pointlike as both point and extended sources. The data was
simulated using \gtobssim\footnote{\gtobssim is distributed
by the Fermi Science Support Center (\cite{fssc}).}
The point sources were simulated with a power-law spectral model with
six $>100$ \mev fluxes ranging from $3\times 10^{-9} \ph/\cm^2/\sec$ to
$10^{-6} \ph/\cm^2/\sec$.  The point sources were simulated with spectral
indices ($\gamma$) of 1.5, 2, 2.5, and 3.  These values were picked to
to represent typical source paramters in LAT-detected sources. The point
sources were simulated on top of an isotropic power-law background
with a Sreekumar-like spectrum ($>100$ \mev flux of $1.5\times 10^{-5}$ and spectral
index 2.1) (\cite{sreekumar_isotropic}).  The Monte Carlo simulation was
performed over a one year time interval using a default rocking profile
and a representative livetime fraction of 0.8.  The reconstruction was
performed with photons from an energy range of 1 \gev to 100 \gev using 4
energy bins per decade and using the 
Pass 7\_V6 (P7\_V6) Source Instrument
Response Function (IRFs).  For each point source, we used \pointlike
to fit it as both a point and extended source and calculate \tsext and
we only kept the sources which had a significant point source detection
($\ts>25$).

@@

We performed a second study to determine the LAT's detection
threshold to spatially extended sources. The detection threshold
to extension is defined as the flux at which the average
value of $\tsext$ is $\langle\tsext\rangle=25$. As was seen in
section~\ref{monte_carlo_validation}, this corresponds to an extension
significance of $5\sigma$.  Qualitatively, we expect the LAT would not
be sensitive to sources much smaller than the PSF or much larger than
the PSF.  Small sources would look too much like the PSF and larger
sources would get lost in the background due to low surface brightness.

To calculate the LAT's detection threshold to spatially extended
sources, we simulated sources with a uniform radially symmetric disk
spatial model on top of a power-law Sreekumar-like isotropic spectrum
(\cite{sreekumar_isotropic}).  The Monte Carlo simulation was performed
using the same one year time interval as in section~\ref{monte_carlo_validation}.
For each extension and spectral index, we
first picked a flux range which bracketed $\tsext=25$. We then picked
ten flux points in this range and repeatedly simulated and performed
extension tests of the extended sources.  For each flux, we performed
$>15$ independent simulations and extension fits.  We calculated
$\langle\tsext\rangle=25$ by fitting a line to all the flux and
$\tsext$ values and estimating the flux when $\tsext=25$.

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extension varying from $\sigma=0.1\deg$
to $2.0\deg$.  The curves on the plot represent the flux for a given
extension where $\langle\tsext\rangle=25$. 
This results shows that the LAT's threshold to extension is better for
harder sources with more high energy photons. This result also shows
that the threshold is best for sources with an extension of $~0.5\deg$
and is significantly worse for very small extended small.

Figure~\ref{index_sensitivity} shows the threshold using photons with
energies between 100 \mev and 100 \gev and also only using photons
with energies between 1 \gev and 100 \gev.  Except for the hardest
and largest sources, our detection threshold is almost unchanged when
including photons with energies between 100 \mev and 1 \gev.  This is also
demonstrated in figure~\ref{ts_ext_vs_energy} which shows \tsext for the
SNR IC443 computed independently in twelve energy bins between 100 \mev and
100 \gev. For IC443, which has a spectral index ~2.4, almost the entire
increase in \ts comes from energies above 1 \gev.  On the other hand,
other systematic errors become increasingly important at low energy. For
our extension search, we use only only photons with energy above 1 \gev.

Figure~\ref{diff_factor_sensitivity} shows how the LAT's detection
threshold varies with increasing diffuse background. We compute the
sensitivity for a power-law source of spectral index 2 using 1, 10, and
100 times the extragalactic background. The latter factor approximately
represents the intensity of the Galactic diffuse emission near the
Galactic center.  The figure shows two plots, the first showing the
threshold using only photons with energies between 1 \gev and 100 \gev and
the second showing the threshold using only photons with energies between
10 \gev and 100 \gev. This plot demonstrates that we are less sensitive
to extended sources in a higher background region and that if the flux of
a source is constant (same number of photons), we would be more sensitive
to the extension of that source if those photons had a higher energy.

@@

We also performed a search for extended sources using photons only between
10 \gev and 100 \gev. Even though we are testing the same sources, this
approach is complimentary because the Galactic diffuse emission is even
less dominant above 10 \gev. Furthermore, source confusion becomes less of
a problem since most LAT sources are not significant at energies above 10
\gev.  Using higher energies, we might expect to detect harder sources in
more complicated regions. This is especially beneficial
for regions near pulsars which cutoff before 10 \gev
and was successfully used in detecting HESS\,J1825-137
and MSH 15-52 with the LAT (\cite{msh1552,fermi_hess_j1825}).

@@

Since extragalactic sources are typically too far away to be resolvable,
we expect most extended sources to be inside our Galaxy and most
likely on the Galactic plane.  Unfortunately, the \gev emission
in the Galactic plane is dominated by extremely structured Galactic
diffuse emission.  Finding sources on top of this emission is difficult
(\cite{first_diffuse_paper}) as has been discussed in 1FGL and 2FGL
(\cite{first_cat,second_cat}).  Furthermore, the Galactic plane is crowded
so it is often difficult to model nearby sources.  Because of this,
finding a source with $\tsext>16$ is not a sufficient for finding
an extended source and we have to perform several analysis crosschecks.

% 2FGL J1856.2+0450c = P72Y3047 for
As an example of why this is important, we describe the test of 2FGL
J1856.2+0450c for extension.  Figure~\ref{example_bad_fit} shows
a smoothed counts map using photons with energy above 1 \gev. In
the map, the Galactic and isotropic diffuse emission are subtracted.
A 2FGL source 2FGL\,J1856.2+0450c was found at the position
$(l,b)=(37.25\deg,0.84\deg)$.  We find that there appears to be overall
some large-scale residual in the diffuse emission along the Galactic
plane. As a result, 2FGL\,J1856.2+0450c fits to a
large extension of 1.83\deg and the result is statistically significant
with \tsext=103.6 (using \pointlike).  By looking at the residuals, it is clear that this
complicated region is not well fit even though the source's extension is
statistically significant. For this reason, we have to perform a series
of further analysis for all of the sources that are found with \tsext=16.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml

First, we took the spatial and spectral model found using \pointlike
and performed a crosscheck by refitting the spectral parameters using
\gtlike.  We used a square ROI $14.1\deg$ in diameter with $1/16\deg$
pixels and the same energy binning as \pointlike. Using \gtlike, we found a second measure of \tsext,
this time not fitting the extension but using a fixed template
with extension determined by \pointlike.
We only considered a source to be significantly extended if \pointlike
and \gtlike both agreed.  For the new extended source candidates, 
table~\ref{alt_diff_model_results} shows a comparison of their the value
of \tsext found by \gtlike and \pointlike.  Everywhere else, the values
from \gtlike are used.

Additionally, we generated for each extended source candidate residual
test statistic (\ts) maps.  The residual \ts map is the \ts value
evaluated at each grid point by adding a new source of spectral index
2 and finding the increase in likelihood when fitting its flux. \ts
maps are especially useful when looking for new sources or other
residual features that are not included in the model. The downside
is they are biased to look for residual features that look like
the PSF and so are not as sensitive to large-scale diffuse features.
features. Furthermore, by definition $\ts>0$ so \ts maps only show
positive fluctuations. Figure~\ref{res_tsmaps} shows an example
residual \ts map for the extended source IC443.

For each candidate, we also generated counts slices. A counts slice
plots the sum of all counts in the longitudinal direction within 2\deg
in the latitudinal direction.  The same plot is also shown integrating
in the longitudinal direction.   An example counts slice is shows for the extended source
IC443 in figure~\ref{counts_slice}. 
On top of the plot is overlaid the model
predicted counts assuming the source has a disk spatial model, assuming
the source is a point source, and fitting just the diffuse emission in
the region. 
Counts slices were especially useful
for looking at how well the overall diffuse emission model fits the
region near the sources.

For each candidate, we also generated a radial integral of the data. This
plot shows the sum of all counts within a given distance from the
source's center. The plot is binned uniformly in $\Delta \theta^2$ so
that an isotropic background will be flat in the plot. On top of this
plot is overlaid the model predicted counts assuming the source has a
disk spatial model, assuming the source is a point source, and fitting
just the diffuse emission in the region. An example radial integral
is shown for the extended source IC443 in figure~\ref{radial_profile}.
This plot is best for comparing the observed counts to the PSF.

For each source, we make a Galactic and extragalactic diffuse emisison
subtracted smoothed counts map. Figure~\ref{smoothed_counts} shows an
example smoothed counts map for IC443. They are useful for qualitativly
assesing the background model.

Finally, for each source, we performed the dual localization method
described in section~\ref{dual_localization_method}.  We did so by
replacing the source with two sources each containing half of the original
flux and then using \pointlike to simultaneously fit the position and
spectrum of these sources.  We then used \gtlike to refit the spectrum of
these sources using the best fit positions. For the new extended sources
found, we summarize the results of this dual localization test and compare
the values of \tsext to \tsinc in table~\ref{dual_localization_results}.

Because of the high source density in the Galactic plane for promising
extended source candidates we often had to iteratively improve the model
of the background sources to obtain a good fit of the extended source.
Often, several catalog sources would fill in the emission of one
extended source and all but one of them would have to be removed from
the model. Similarly, the position of background sources would often
have to be refit once the extended source's extension was correctly
fit. Similarly, background sources were insignificant in the fit
energy range and had to be removed.  When the position and extension
of the sources couples strongly with the bacground catalog sources, we
iteratively fit the position and extended of all nearby sources until the
fit stabilized.  For each extended source, we describe what modifications
were required to the sky model to obtain a good fit of the region.

\subsection{Alternative Diffuse Model Crosscheck}
\label{alt_diff_model_description}

To estimate systematic errors due to uncertainty in the diffuse
emission background model, we performed a second
extension test of the candidates using
an alternative diffuse
model. The alternative model was based on the GALPROP\footnote{GALPROP
is a software package for calculating the Galactic $\gamma$-ray
emission based on a model of cosmic-ray propagation in the Galaxy. See
\url{http://galprop.stanford.edu/} for details and references} code
used in the LAT analysis of the isotropic diffuse emission
(\cite{isotropic_lat}).  
With this model,
we independently fit in the nearby region 
7 HI rings, 7 HII rings, the inverse component,
the E(B-V) residual, and the isotropic component of
the background ration.
All told, we fit in the region 18 individual templates along
with all source parameters during the analysis.

>>>>

\usepackage[textwidth=0.8in,colorinlistoftodos]{todonotes}
\usepackage[toc,page]{appendix}

@@

% some units
\newcommand{\kev}{\text{Kev}\xspace}
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\renewcommand{\sec}{\text{s}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_{\text{ext}}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_{\text{inc}}}\xspace}
\newcommand{\loglikelihood}{\ensuremath{LL}\xspace}

@@

\newcommand{\tsextpointlike}{\ensuremath{\tsext_{,\pointlike}}\xspace}
\newcommand{\tsextgtlike}{\ensuremath{\tsext_{,\gtlike}}\xspace}

\newcommand{\ts}{\text{TS}\xspace}
\newcommand{\glon}{\text{GLON}\xspace}
\newcommand{\glat}{\text{GLAT}\xspace}
\newcommand{\altdiff}{\text{alt,diff}\xspace}
\newcommand{\altpsf}{\text{alt,psf}\xspace}
\newcommand{\sys}{\text{sys}\xspace}
\newcommand{\stat}{\text{stat}\xspace}

@@

% the program names
\newcommand{\pointlike}{\text{\em pointlike}\xspace}
\newcommand{\python}{\text{\em python}\xspace}
\newcommand{\gtlike}{\text{\em gtlike}\xspace}
\newcommand{\gtobssim}{\text{\em gtobssim}\xspace}
\newcommand{\minuit}{\text{\em Minuit}\xspace}
\renewcommand{\approx}{\sim\!\xspace}

@@

The Large Area Telescope (LAT) is a pair conversion telescope on the
Fermi Gamma-Ray Space Telescope (Fermi). It has been surveying the
$\gamma$-ray sky since June 2008.  Fermi is well suited as a survey
instrument because of its broad energy coverage (20 \mev to $>300$
\gev), wide field of view ($\approx 2.4 \sr$), and large effective area
($\approx 8000 \cm^2$ at $>1 \gev$).

Using two years of all sky surveying data, the LAT Collaboration
published a catalog of 1873 \gev sources\cite{second_cat}. Many of these can
be positionally associated with a variety of source classes including
Active Galactic Nuclei (AGN), pulsars, galaxies, and supernova
remnants (SNRs).  While many of these classes can be spatially
resolved when observed at other frequencies, detection an extension at
\gev energies by the LAT is complicated by the rather large
point-spread function (PSF) of the instrument (the angular resolution
for events converting in the front part of the detector, defined as
the 68\% containment radius for a point source, is approximately
$3.5^{\circ}$ at 100 \mev, improving to about 0.1$^{\circ}$ at 10 \gev).

@@

In addition to these \gev $\gamma$-ray detections, a variety of
spatially extended sources have been detected in the Galactic plane at
\tev energies using Air Cherenkov detectors. Most prominent was a
survey of the Galactic plane using the High Energy Stereoscopic System
(H.E.S.S) which presented 14 spatially extended sources
(\cite{hess_plane_survey}) sources with extensions varying from
$~0.1\deg$ to $~0.25\deg$. In fact, within our Galaxy only very few
sources (most notably the $\gamma$-ray binaries
LS\,5039 (\cite{HESSLS5039}), LS I+61-303 (\cite{MAGICLSI, VERITASLSI}) and
HESS\,J0632+057 (\cite{HESS0632}) are amongst the few Galactic sources
for which no significant extension was detected at \tev
energies. Given these findings, it can be expected that if the source
population at \gev energies has similar properties, that at least some
of the Fermi-LAT sources in our Galaxy should be spatially extended at
\gev energies.

Being able to spatially resolve the \gev emission from a source is
important for several reasons. Because of the large PSF and the large
candidate source density in our Galaxy, source confusion is a severe
problem for the Fermi-LAT. Finding a matching morphology can be used
to uniquely identify the source.  Because of the strongly varying PSF
of the LAT, the spatial and spectral information about a source do not
simply decouple. Therefore, a biased spatial model will bias the
spectral model of the source and studying the extension of the source
can correct for this.  Furthermore, correctly modeling a source's
extension is important for improving the model of the sky and
removing excess residuals.  These residuals manifested themselves in
1FGL as spurious point sources in the region region surrounding the
Large Magellanic Cloud (\cite{first_cat}).  For this reason, we are
motivated to take the 1873 sources that were detected in 2FGL and
apply a search for those sources which are spatially extended.

@@

The convolution of radially symmetric extended sources is optimized by
using a semi-analytic calculation.  The observed photon distribution
can be written as the convolution of the source shape ($I_\text{src}$)
with the PSF
\begin{equation}
  \text{PDF}(\vec r) = \int  \text{PSF}(|\vec r - \vec r'|)I_\text{src}(\vec r') d A'.
\end{equation}
For the LAT, the PSF is parameterized by a {\em King function}
\begin{equation}
  \text{PSF}(\vec r) = 
  \frac{1}{2\pi\sigma^2}
  \left(1-\frac{1}{\gamma}\right)
  \left(1+\frac{u}{\gamma}\right)^{-\gamma},
\end{equation}
where $\sigma$ and $\gamma$ are free parameters
(\cite{matthew_kerr_thesis}).  For radially symmetric extended sources,
the angular part of the integral can be done analytically
\begin{equation}
  \text{PDF}(u)= \int_0^\infty dv
  I_\text{src}(v) 
  \left(\frac{\gamma-1}{\gamma}\right)
  \left( \frac{\gamma}{\gamma + u + v}\right)^\gamma 
  \times ~_2F_1 \left(\gamma/2,\frac{1+\gamma}{2},1,\frac{4uv}{(\gamma+u+v)^2}\right).
\end{equation}
This convolution formula is used by \pointlike because it is faster
and more closely reduces to the PSF when the source's size is 
small.

@@

We found that the {\em HESSE} errors on the extension of a source
were sometimes badly estimated.  On the other hand, the {\em Simplex}
algorithm was prohibitively slow and would often fail to convergence. Our
approach to calculating extension errors is to fix the position of the
source and varying just the extension using the {\em Simplex} algorithm.
This approach makes the approximation that the covariance between each
spatial parameter is small but not that the likelihood is Gaussian. This
method is shown schematically in figure~\ref{four_plots_ic443}(a) which
shows the change in \loglikelihood when varying the extension of the
SNR IC443.  The localization error is separately calculated by fixing
the extension and spectrum of the source and fitting a parabola to the
likelihood as a function of position.

@@

The half $\chi^2_1$ distribution was found to hold for the special case
of source finding using EGRET (\cite{mattox_egret}). This distribution
is also plausible when testing point sources for extension. Due to
statistical fluctuations, half of the point sources might look narrower
than the PSF leading to $\tsext=0$. The other half of the time, we would
get a distribution consistent with Wilks' theorem. We will present a
Monte Carlo study that shows an empirical distribution of \tsext in the
null hypothesis and compares it to this distribution.

We simulated point sources of varying spectral models and fit them
them using \pointlike as both point and extended sources. The data was
simulated using \gtobssim\footnote{\gtobssim is distributed
by the Fermi Science Support Center (\cite{fssc}).}
The point sources were simulated with a power-law spectral model with
six $>100$ \mev fluxes ranging from $3\times 10^{-9} \ph/\cm^2/\sec$ to
$10^{-6} \ph/\cm^2/\sec$.  The point sources were simulated with spectral
indices ($\gamma$) of 1.5, 2, 2.5, and 3.  These values were picked to
to represent typical source parameters in LAT-detected sources. The point
sources were simulated on top of an isotropic power-law background
with a Sreekumar-like spectrum ($>100$ \mev flux of $1.5\times 10^{-5}$ and spectral
index 2.1) (\cite{sreekumar_isotropic}).  The Monte Carlo simulation was
performed over a one year time interval using a default rocking profile
and a representative livetime fraction of 0.8.  The reconstruction was
performed with photons from an energy range of 1 \gev to 100 \gev using 4
energy bins per decade and using the 
Pass 7\_V6 (P7\_V6) Source Instrument
Response Function (IRFs).  For each point source, we used \pointlike
to fit it as both a point and extended source and calculate \tsext and
we only kept the sources which had a significant point source detection
($\ts>25$).

@@

We calculated the LAT's detection threshold to spatially
extended sources. The detection threshold to extension is
defined as the flux at which the average value of $\tsext$ is
$\langle\tsext\rangle=16$ corresponding to a $4\sigma$ detection (see
section~\ref{monte_carlo_validation}).  Qualitatively, we expect the LAT
would not be sensitive to sources much smaller than the PSF or much larger
than the PSF.  Small sources would look too much like the PSF and larger
sources would get lost in the background due to low surface brightness.

To calculate the LAT's detection threshold, we simulated sources with
a uniform radially symmetric disk spatial model on top of a power-law
Sreekumar-like isotropic spectrum (\cite{sreekumar_isotropic}).  The Monte
Carlo simulation was performed using the same two year time range that
was used in 2FGL.  For each extension and spectral index, we picked a
flux range which bracketed $\tsext=16$. We then picked ten flux points in
this range and repeatedly simulated and then performed extension tests of
the simulated extended source.  We calculated $\langle\tsext\rangle=16$
by fitting a line to the flux and $\tsext$ values.

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extension varying from $\sigma=0.1\deg$
to $2.0\deg$.  The LAT's threshold to extension
is better for harder sources with more high energy photons and
the threshold is best for sources with an extension 
$\approx0.5\deg$ and is significantly worse for very small extended small.
Figure~\ref{index_sensitivity} shows the threshold using photons with
energies between 100 \mev and 100 \gev and also using only photons
with energies between 1 \gev and 100 \gev.  Except for the hardest
and largest sources, our detection threshold is almost unchanged when
including photons with energies between 100 \mev and 1 \gev.  This is also
demonstrated in figure~\ref{four_plots_ic443}(b) which shows \tsext for the
SNR IC443 computed independently in twelve energy bins between 100 \mev and
100 \gev. For IC443, which has a spectral index $\approx2.4$, almost the entire
increase in \ts comes from energies above 1 \gev.  On the other hand,
other systematic errors become increasingly important at low energy. For
our extension search, we use only only photons with energy above 1 \gev.

Figure~\ref{diff_factor_sensitivity} shows how the LAT's detection
threshold increases with increasing background. We compute the threshold
using 1, 10, and 100 times the Sreekumar-like isotropic background. The
latter factor approximately represents the intensity of the Galactic
diffuse emission near the Galactic center.  The figure shows the threshold
using only photons with energies between 1 \gev and 100 \gev and aldo
when using only photons with energies between 10 \gev and 100 \gev. We
are less sensitive to extended sources with higher background. Also,
given the same flux we are more sensitive to the source if those
photons have a higher energies.

Overlaid on figure~\ref{diff_factor_sensitivity} are the LAT
detected extended sources from table~\ref{known_extended_sources} and
table~\ref{new_ext_srcs_table}.  The extended sources are color coded
based upon whether the galactic diffuse and isotropic emission model is
closer to 1, 10, or 100 times the Sreekumar-like isotropic spectrum.

Finally, figure~\ref{all_sensitivity} shows the LAT's detection threshold
for all four spectral indices (1.5, 2, 2.5, and 3) against all three
background levels (1, 10, and 100 times the Sreekumar-like isotropic
background).  For a fixed 1 \gev to 100 \gev or 10 \gev to 100 \gev
flux, the LAT's detection threshold is only weakly depends upon the
spectral index.  This effect is most pronounced when using only photons
with energies above 10 \gev. The extension thresholds are tabulated in
table~\ref{all_sensitivity_table}.

@@

We also performed a search for extended sources using photons only
between 10 \gev and 100 \gev. Even though we are testing the same
sources, this approach is complimentary because the Galactic diffuse
emission is even less dominant above 10 \gev. Furthermore, source
confusion becomes less of a problem since most LAT sources are not
significant at energies above 10 \gev.  Using higher energies, we might
expect to detect harder sources in more complicated regions. This is
beneficial for regions near pulsars which cutoff before 10 \gev and was
successfully used in detecting HESS\,J1825-137 and MSH 15-52 with the LAT
(\cite{msh1552,fermi_hess_j1825}).

@@

Since extragalactic sources are typically too far away to be resolvable,
we expect most extended sources to be inside our Galaxy and most
likely on the Galactic plane.  Unfortunately, the \gev emission
in the Galactic plane is dominated by extremely structured Galactic
diffuse emission.  Finding sources on top of this emission is difficult
(\cite{first_diffuse_paper}) as has been discussed in 1FGL and 2FGL
(\cite{first_cat,second_cat}).  Furthermore, the Galactic plane is crowded
and it is often difficult to correctly model nearby sources.  Because of this,
finding a source with $\tsext>16$ is not a sufficient
criteria for resolving a source. For each extended source,
we perform several analysis crosschecks.

% 2FGL J1856.2+0450c = P72Y3047 for
As an example of why this is important, we describe the test of 
2FGL\,J1856.2+0450c for extension.  Figure~\ref{example_bad_fit} shows
a Galactic and isotropic diffuse emission subtracted
smoothed counts map of this source using photons with energy 
between 1 \gev and 100 \gev. 
In this region along the Galactic plane, there appears to be 
large-scale residual in the diffuse emission. As a result, 2FGL\,J1856.2+0450c 
fit to an extension of 1.83\deg and the result is statistically significant
with \tsext=45.4.  By looking at the residuals, it is clear that this
complicated region is not fit well even though the source's extension is
statistically significant. For this reason, we manually discard
the extension of this and similar sources.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml

For each extended source candidate,
we took the spatial and spectral model found using \pointlike and
performed a crosscheck by refitting the spectral parameters using
\gtlike.  We used a square ROI $14.1\deg$ in diameter with $1/32\deg$
spatial bins and the same energy binning as \pointlike. Using \gtlike,
we found a second measure of \tsext, this time not fitting the extension
but using a fixed template with the extension determined by \pointlike.
We only considered a source to be significantly extended if \pointlike
and \gtlike both agreed.  For the new extended source candidates,
table~\ref{alt_diff_model_results} will show a comparison of of \tsext for
\gtlike and \pointlike.

For each candidate, we generated a map of residual \ts
by adding a new source of spectral index 2 into the region and finding
the increase in likelihood when fitting its flux. \ts maps are useful
to look for residual structure in the sky mode.  Figure~\ref{res_tsmaps}
shows a residual \ts map for the extended source IC443.

For each candidate, we generated plots of the sum of all counts
within a given distance of the source and compared it to the model
predictions of a point source.  The plots were binned uniformly
in $\Delta \theta^2$ so that an isotropic background would be flat.
An example radial integral plot is shown for the extended source IC443
in figure~\ref{four_plots_ic443}(c).  For each source, we also made
Galactic and extragalactic diffuse emission subtracted smoothed counts
maps. Figure~\ref{four_plots_ic443}(d) shows an example for IC443.

For each source, we tested to see if the region contained instead
multiple point sources by replacing the extended source with
two point sources and performed the dual localization procedure
described in section~\ref{dual_localization_method}.  We only
considered a source to be extended if $\tsext>\tsinc$.  For the new
extended sources, we will show the results of this dual localization in
table~\ref{dual_localization_results}.

Because of the high source density in the Galactic plane, for promising
candidates we often had to iteratively improve the model of our background
sources to obtain a better fit of the sources.  Often, several catalog
sources would fill in the emission of the extended source and would
have to be removed from the region.  Similarly, background sources
were often insignificant in the fit energy range and had to be removed.
The position of background sources would often have to be refit once the
extended source was fit.  When the model of the extended source coupled
strongly with nearby sources, we iteratively fit the extended source and
all nearby sources until the fit converged.  For each extended source,
we describe the modification that were required.

\subsection{Systematic Errors on Extension}
\label{systematic_errors_on_extension}

% Alternative PSF

We estimate a systematic error on the extension of a source due to
uncertainty in our knowledge of the angular resolution of the LAT.  Before
launch, the LAT PSF was determined by a detector simulation which was
verified in accelerator tests (\cite{atwood_LAT_mission}).  A discrepancy
was found in flight data above a few \gev in the PSF compared to the
shape of bright AGN.  A publication on this issue is in preparation
(\cite{lat_on_orbit_psf}).  Subsequently, the PSF was fit empirically to
bright AGN and this empirical parameterization is the default PSF used
in the P7\_V6 IRFs.  To account for this uncertainty in our knowledge of
the PSF, we refit our extended source candidates using the Monte Carlo
representation of the PSF and consider the difference in extension found
using the two PSFs as a systematic error on the extension of a source.
Because the Monte Carlo representation of the PSF is smaller, it leads
to unphysically higher \tsext values.  This same approach was used in
(\cite{ic443}).  At high energies, we believe that our parameterization
of the PSF from bright AGN is substantially better than the Monte Carlo
representation of the PSF so this systematic errors is conservative.

% Alternate Diffuse

We estimate a second systematic error on the extension of a source
due to uncertainty in the model of the Galactic diffuse emission by
using an alternative diffuse model. The alternative model was based on
the GALPROP\footnote{GALPROP is a software package for calculating the
Galactic $\gamma$-ray emission based on a model of cosmic-ray propagation
in the Galaxy. See \url{http://galprop.stanford.edu/} for details and
references} model used in the LAT analysis of the isotropic diffuse
emission (\cite{isotropic_lat}).  With this model, we independently fit
along with the point and diffuse sources in the region seven HI rings,
seven HII rings, the inverse component, the E(B-V) residual, and the
isotropic component of the background emission.  All told, we fit in
the region 18 diffuse templates.  This same crosscheck was used in the
LAT Collaboration's analysis of RX J1713.7-3946 (\cite{rx_j1713_lat}).

It is not expected that this diffuse model is more physical.
But adding degrees of freedom to the background model can remove
likely spurious sources that correlate with features in the Galactic
diffuse emission.  Therefore, this tests systematics that may be
due to incorrect modeling of the diffuse emission in the region.
The results of this fit for the new extended source candidates are shown
in table~\ref{alt_diff_model_results}. There is good agreement between
the \ts using the standard and alternative diffuse models, suggesting
that the sources are robust against features in the diffuse model.
